Automated Residual Plot Assessment with the R Package autovi and Shiny App autovi.web

### 2024-11-18 autovi_paper.qmd
drastic paper trimming

title: "Software for Automated Residual Plot Assessment: autovi and autovi.web"

@@

    shorttitle: "Software for Automated Residual Plot Assessment"

@@

The main purpose of `autovi` is to provide rejection decisions and $p$-values for testing whether a regression model is correctly specified. The package introduces a novel approach to automating statistical analysis, particularly in the interpretation of residual plots. The name `autovi` stands for automated visual inference. <!--While initially designed for linear regression residual diagnostics, it has the potential to be extended to broader visual inference applications, as we discuss in section @sec-autovi-infrastructure.-->

@@

```{r}
#| label: three-examples
#| fig-cap: "Reading residual plots can be a difficult task, particular for students new to statistical modeling. The autovi package makes it easier. Here are three examples of residual plots, which may appear to have structure. According to autovi, the visual signal strengths (VSS) of these three examples are approximately (a) 1.53, (b) 3.57, (c) 5.87, resulting in (b), (c) being significant violations of good residuals, but (a) is consistent with a good residual plot."
#| fig-width: 9
#| fig-height: 3
#| out-width: 100%
library(readr)
library(dplyr)
library(ggplot2)
library(patchwork)
library(ggthemes)

@@

The `autovi` package is available on CRAN. It is actively developed and maintained, with the latest updates accessible on GitHub at [https://github.com/TengMCing/autovi](https://github.com/TengMCing/autovi). The code discussed in this paper is based on `autovi` version 0.4.1.

@@

## Usage {#sec-autovi-usage}

To get started quickly, only three function calls are needed to obtain a summary of the automated residual assessment:

```{r}
set.seed(10086)
```

@@

@@

The three functions are explained as follows:

The summary reports key findings such as the VSS of the true residual plot and the $p$-value of the automated visual test. The $p$-value is the ratio of null plots that have VSS greater than or equal to that of the true residual plot. We typically reject the null hypothesis when the $p$-value is smaller than or equal to a desired significance level, such as 5%. The report also provides sample quantiles of VSS for null and bootstrapped plots, helping to explain the severity and likelihood of model violations.

Although the $p$-value is sufficient for automated decision-making, users can visually inspect the original residual plot alongside a sample null plot. This visual comparison can clarify why $H_0$ is either rejected or not, and help identify potential remedies. The `plot_pair()` and `plot_lineup()` methods facilitate this comparison.

```{r fig-plot-pair, echo = TRUE, fig.height=3, fig.width=4, fig.cap="True plot alongside one null plot, for quick comparison."}
checker$plot_pair()
```

@@

The modules for VSS prediction and $p$-value computation are predefined and cannot be overridden, although users can interact with them directly through function arguments. Similarly, the image reading and resizing module is fixed but will adapt to different Keras models by checking their input shapes. The remaining seven modules are designed to be overridable, enabling users to tailor the infrastructure to their specific needs. These modules are discussed in detail in the following sections.

### Initialization

@@

The `get_keras_model()` function can be used to download a model to a temporary directory and load it into memory using `TensorFlow`. It requires only the model name, which is the value in the first column of the `tibble` returned by `list_keras_model()`.

@@

The `autovi.web` package extends the functionality of `autovi` by offering a user-friendly web interface for automated residual plot assessment. This eliminates the common challenges associated with software installation, so users can avoid managing Python environments or handling version requirements for R libraries. The platform is cross-platform and accessible on various devices and operating systems, making it suitable even for users without R programming experience. Additionally, updates are managed centrally, ensuring that users always have access to the latest features.

The `autovi.web` interface is available at [autoviweb.netlify.app](autoviweb.netlify.app). This section discusses the implementation based on `autovi.web` version 0.1.0.

>>>>

title: "Automated Residual Plot Assessment with the R Package autovi and Shiny App autovi.web"

@@

    shorttitle: "Automated Residual Plot Assessment"

@@

The main purpose of `autovi` is to provide rejection decisions and $p$-values for testing the null hypothesis ($H_0$) that the regression model is correctly specified. The package provides automated interpretation of residual plots using computer vision. The name `autovi` stands for **auto**mated **v**isual **i**nference. <!--While initially designed for linear regression residual diagnostics, it has the potential to be extended to broader visual inference applications, as we discuss in section @sec-autovi-infrastructure.-->

@@

```{r}
#| label: fig-three-examples
#| fig-cap: "Reading residual plots can be a difficult task, particular for students new to statistical modeling. The autovi package makes it easier. Here are three examples of residual plots, which may appear to have structure. According to autovi, the visual signal strengths (VSS) of these three examples are approximately (a) 1.53, (b) 3.57, (c) 5.87, resulting in (b), (c) being significant violations of good residuals, but (a) is consistent with a good residual plot."
#| fig-width: 12
#| fig-height: 4
#| out-width: 100%
library(readr)
library(dplyr)
library(ggplot2)
library(patchwork)
library(ggthemes)

@@

The `autovi` package is available on CRAN. It is actively developed and maintained, with the latest updates accessible on GitHub. The code discussed in this paper is based on `autovi` version 0.4.1.

@@

We recommend using the Shiny app `autovi.web` if encountering installation problems.

## Numerical Summary {#sec-autovi-usage}

Three steps are needed to get an automated assessment of a set of residuals and fitted values:

The code to do this is:

```{r}
#| echo: false
set.seed(10086)
```

@@

It produces the following summary:

@@

The summary includes observed VSS of the true residual plot and associated $p$-value of the automated visual test. The $p$-value is the proportion of null plots (out of the total 100) that have VSS greater than or equal to that of the true residual plot. <!--We typically reject the null hypothesis when the $p$-value is smaller than or equal to a desired significance level, such as 5%.--> The report also provides sample quantiles of VSS for null samples and bootstrapped data plots, providing more information about the sampling variability and a likelihood of model violations. The likelihood is computed from the proportion of values greater than the observed VSS in both the bootstrapped data values and the simulated null values.

## Visual Summary {#sec-autovi-visual}

Users can visually inspect the original residual plot alongside a sample null plot using `plot_pair()` or a lineup of null plot `plot_lineup()`. This visual comparison can clarify why $H_0$ is either rejected or not, and help identify potential remedies.

```{r fig-plot-pair, echo = TRUE, fig.height=2, fig.width=3, out.width="60%", fig.cap="True plot alongside one null plot, for quick comparison."}
checker$plot_pair()
```

@@

The modules for VSS prediction and $p$-value computation are predefined and cannot be overridden, although users can interact with them directly through function arguments. Similarly, the image reading and resizing module is fixed but will adapt to different Keras models by checking their input shapes. The remaining seven modules are designed to be overridable, enabling users to tailor the infrastructure to their specific needs. These modules are discussed in detail on the software's website. <!-- in the following sections.-->

<!--
### Initialization

@@

The `get_keras_model()` function can be used to download a model to a temporary directory and load it into memory using `TensorFlow`. It requires only the model name, which is the value in the first column of the `tibble` returned by `list_keras_model()`.
-->

@@

The `autovi.web` package extends the functionality of `autovi` by offering a user-friendly web interface for automated residual plot assessment. This eliminates the common challenges associated with software installation, so users can avoid managing Python environments or handling version requirements for R libraries. The platform is cross-platform and accessible on various devices and operating systems, making it suitable even for users without R programming experience. Additionally, updates are managed centrally, ensuring that users always have access to the latest features.
<!--The `autovi.web` interface is available at [autoviweb.netlify.app](autoviweb.netlify.app).--> This section discusses the implementation based on `autovi.web` version 0.1.0.
